Wikipedia WALS Feature Analysis

### 2019-07-19 surprisal/WALS_features.Rmd
most features imputed for all languages

```{r setup, include=FALSE}
library(tidyverse)
library(lingtypology)
library(feather)
library(here)
library(missMDA)
library(lsa)
library(reshape2)
library(broom)
library(janitor)
knitr::opts_chunk$set(echo = TRUE)
```

@@

```{r feature space}
imputed_df <- read_csv(here("Data/imputed_wals.csv"))

feature_space <- imputed_df %>% 
  gather(feature, value, -language) %>% 
  group_by(feature) %>% 
  mutate(value = as.numeric(as.factor(value))) %>% 
  spread(feature, value) %>%
  column_to_rownames("language") %>%
  dist(method = "manhattan") %>%
  as.matrix() %>%
  melt(value.name = "feature") %>%
  as_tibble() %>%
  rename(language1 = Var1, language2 = Var2) 
```

```{r cosines}
common_languages <- read_csv(here("Data/imputed_wals.csv")) %>% 
  pull(language)

unigram_cosines <- wiki %>%
  filter(language %in% common_languages, gram == "Unigram") %>%
  select(-gram) %>% 
  column_to_rownames("language") %>%
  t() %>%
  cosine() %>%
  as_data_frame(rownames = "language1") %>%
  gather(language2, unigram_cosine, -language1) %>%
  mutate(unigram_cosine = 1 - abs(unigram_cosine))

bigram_cosines <- wiki %>%
  filter(language %in% common_languages, gram == "Bigram") %>%
  select(-gram) %>% 
  column_to_rownames("language") %>%
  t() %>%
  cosine() %>%
  as_data_frame(rownames = "language1") %>%
  gather(language2, bigram_cosine, -language1) %>%
  mutate(bigram_cosine = 1 - abs(bigram_cosine))

trigram_cosines <- wiki %>%
  filter(language %in% common_languages, gram == "Trigram") %>%
  select(-gram) %>% 
  column_to_rownames("language") %>%
  t() %>%
  cosine() %>%
  as_data_frame(rownames = "language1") %>%
  gather(language2, trigram_cosine, -language1) %>%
  mutate(trigram_cosine = 1 - abs(trigram_cosine))

all_distances <- feature_space %>%
  left_join(unigram_cosines) %>%
  left_join(bigram_cosines) %>%
  left_join(trigram_cosines)

all_distances %>% 
  select(-bigram_cosine) %>% 
  ggplot(aes(x = feature)) +
    geom_smooth(aes(y = unigram_cosine, color = "#0033CC"), 
                method = 'lm', se = F) + 
    # geom_smooth(aes(y = bigram_cosine, color = "#6699FF"), 
    #             method = 'lm', se = F) +
    geom_smooth(aes(y = trigram_cosine, color = "#33CCFF"), 
                method = 'lm', se = F) +
    geom_point(aes(y = unigram_cosine, color = "#0033CC")) + 
    # geom_point(aes(y = bigram_cosine, color = "#6699FF")) + 
    geom_point(aes(y = trigram_cosine, color = "#33CCFF")) + 
    ylab("Information Curve Difference") + 
    xlab("Difference in WALS Features") + 
    scale_color_manual(name = "Gram", 
                       values = c("#0033CC", "#33CCFF"), 
                       labels = c("Unigrams", "Trigrams")) 
```

```{r models}
imputed_df <- read_csv(here("Data/imputed_wals.csv"))

wf_sub <- imputed_df %>% 
  group_by(language) %>%
  slice(1) %>%
  gather(feature, value, -language) %>%
  group_by(feature) %>%
  mutate(value = as.numeric(as.factor(value)))

@@

```{r imputation}
imputed_df <- common_df %>% 
  gather(feature, value, -language) %>%
  group_by(feature) %>%
  mutate(value = factor(value)) %>%
  select(language, feature, value) %>%
  ungroup() %>% 
  spread(feature, value) %>% 
  sapply(as.factor) %>%
  as.data.frame() %>%
  column_to_rownames("language") %>% 
  MIMCA(ncp = 1)

imputed_df <- as.data.frame(imputed_df[[1]][[99]]) 
```

@@

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(lingtypology)
library(feather)
library(here)
library(missMDA)
library(lsa)
library(reshape2)
library(broom)
library(janitor)
library(hexbin)
knitr::opts_chunk$set(echo = TRUE)
```

@@

```{r imputation}
df <- read_csv(here("Data/wiki_wals.csv"))

wals <- df %>% 
  select(language, ends_with('A')) %>%
  group_by(language) %>%
  slice(1) %>% 
  ungroup()

imputed_df <- wals %>% 
  select(-`139A`, -`141A`, -`140A`) %>% # these features have no values
  select(language, `121A`:`120A`) %>%
  gather(feature, value, -language) %>%
  group_by(feature) %>%
  mutate(value = factor(value)) %>%
  select(language, feature, value) %>%
  ungroup() %>% 
  spread(feature, value) %>% 
  sapply(as.factor) %>%
  as.data.frame() %>%
  column_to_rownames("language") %>% 
  MIMCA(ncp = 2, threshold = 1e-02)

imputed_df110 <- as.data.frame(imputed_df[[1]][[99]]) 
```

```{r models}
imputed_df <- read_csv(here("Data/imputed_all.csv"))

wf_sub <- all_imputed %>% 
  group_by(language) %>%
  slice(1) %>%
  gather(feature, value, -language) %>%
  group_by(feature) %>%
  mutate(value = as.numeric(as.factor(value)))

@@

```{r cosines}
unigram_cosines <- wiki %>%
  filter(language %in% sub_langs, gram == "Unigram") %>%
  select(-gram) %>% 
  column_to_rownames("language") %>%
  t() %>%
  cosine() %>%
  as_data_frame(rownames = "language1") %>%
  gather(language2, unigram_cosine, -language1) %>%
  mutate(unigram_cosine = 1 - abs(unigram_cosine))

# bigram_cosines <- wiki %>%
#   filter(language %in% sub_langs, gram == "Bigram") %>%
#   select(-gram) %>% 
#   column_to_rownames("language") %>%
#   t() %>%
#   cosine() %>%
#   as_data_frame(rownames = "language1") %>%
#   gather(language2, bigram_cosine, -language1) %>%
#   mutate(bigram_cosine = 1 - abs(bigram_cosine))

trigram_cosines <- wiki %>%
  filter(language %in% sub_langs, gram == "Trigram") %>%
  select(-gram) %>% 
  column_to_rownames("language") %>%
  t() %>%
  cosine() %>%
  as_data_frame(rownames = "language1") %>%
  gather(language2, trigram_cosine, -language1) %>%
  mutate(trigram_cosine = 1 - abs(trigram_cosine))

all_distances <- feature_space %>%
  left_join(unigram_cosines) %>%
  left_join(trigram_cosines)

all_distances %>% 
  ggplot(aes(x = feature)) +
    geom_smooth(aes(y = unigram_cosine, color = "#0033CC"), 
                method = 'lm', se = F) + 
    # geom_smooth(aes(y = bigram_cosine, color = "#6699FF"), 
    #             method = 'lm', se = F) +
    geom_smooth(aes(y = trigram_cosine, color = "#33CCFF"), 
                method = 'lm', se = F) +
    geom_hex(aes(y = unigram_cosine, color = "#0033CC", alpha=.2)) + 
    # geom_point(aes(y = bigram_cosine, color = "#6699FF")) + 
    geom_hex(aes(y = trigram_cosine, color = "#33CCFF", alpha=.2)) + 
    ylab("Information Curve Difference") + 
    xlab("Difference in WALS Features") + 
    scale_color_manual(name = "Gram", 
                       values = c("#0033CC", "#33CCFF"), 
                       labels = c("Unigrams", "Trigrams")) 
```

@@

```{r feature space}
imputed_df <- read_csv(here("Data/imputed_all.csv"))

feature_space <- all_imputed %>% 
  gather(feature, value, -language) %>% 
  group_by(feature) %>% 
  mutate(value = as.numeric(as.factor(value))) %>% 
  spread(feature, value) %>%
  column_to_rownames("language") %>%
  dist(method = "manhattan") %>%
  as.matrix() %>%
  melt(value.name = "feature") %>%
  as_tibble() %>%
  rename(language1 = Var1, language2 = Var2) 
```
